---
title: "Lab 5"
author: "Seungah Lee"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Load Data**

```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(tidyverse)
library(readxl)
library(lubridate)
library(maps)
```


```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```


<br>

1. Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?

```{r}
accday <- acc %>%
  within(DAY_WEEK <- factor(DAY_WEEK, labels = c('Sun', 'Mon', 'Tues', 'Wed', 'Thurs', 'Fri', 'Sat')))
  
ggplot(accday, aes(x=DAY_WEEK)) + 
  geom_bar() +
  labs(title="Number of accidents by day", x="Weekday", y="Count")
```

The most accidents occur on Saturdays, followed by accidents on Sundays and Fridays.

<br>

2. Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)

```{r}
fatal <- person %>%
  filter(INJ_SEV == '4')

head(fatal)
```

3. Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)

```{r}
danger <- fatal %>%
  drop_na(MAKE) %>%
  group_by(STATE, MAKE) %>%
  summarise(Accidents = n()) %>%
  group_by(STATE) %>%
  filter(Accidents == max(Accidents))

head(danger)
```

4. Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)


```{r, message = FALSE, warning = FALSE}
glc <- readxl::read_xlsx("FRPP_GLC_-_United_StatesSep292020.xlsx")
states <- map_data('state')

state_code <- glc %>% mutate('State Code' = as.numeric(`State Code`))

vehicle <- danger %>%
  inner_join(state_code %>% select("State Name", "State Code"), by = c("STATE" = "State Code")) %>% distinct()

vehicle$`State Name` <- tolower(vehicle$"State Name")

perstate <- states %>%
  left_join(vehicle %>% select("State Name", "MAKE"), by = c("region"= "State Name"))

com <- perstate %>% 
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat), make = min(MAKE))

com$make <- factor(com$make, levels = c(12, 20, 37, 49), labels = c("Ford", "Chevrolet", "Honda", "Toyota"))

ggplot(perstate %>% group_by(region), aes(x = long, y = lat)) + 
  geom_path(aes(group=group)) + 
  geom_text(data = com, aes(label = make), size = 3)
```

The most dangerous vehicle is the one with the highest fatality rate.


5. Join the accident and person table (work out which variable(s) to use)

```{r message=FALSE, warning=FALSE}
accper <- acc %>%
  select(c(DAY_WEEK, STATE, HOUR)) %>%
  inner_join(person %>% select(c(STATE, SEX)), by = 'STATE')
```

6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.

```{r, message = FALSE}
acctal <- accper %>%
  group_by(DAY_WEEK, HOUR, SEX) %>%
  summarise(n = n())

acctal %>%
  ggplot(aes(x = HOUR, weight = n, fill = SEX)) +
  geom_histogram() +
  facet_grid(~DAY_WEEK) +
  labs(title="Number of accidents by hour", x="Hour when accidents", y="Total Accidents", fill="Gender")
```

7. Now plot a choropleth map of the number of deaths on a county level.


```{r, message=FALSE}
counties <- map_data("county")

condea <- acc %>% 
  group_by(STATE, COUNTY) %>% 
  summarise(choropleth = sum(FATALS))

glc2 <- glc %>% 
  rename(STATE=`State Code`, COUNTY=`County Code`) %>% 
  mutate(STATE=as.numeric(STATE), COUNTY=as.numeric(COUNTY))

choro <- condea %>% 
  inner_join(glc2, by=c("STATE", "COUNTY")) %>% 
  select(choropleth,`State Name`, `County Name`) %>% 
  distinct() %>%
  rename(region=`State Name`, subregion=`County Name`) %>% 
  mutate(region=tolower(region), subregion=tolower(subregion))%>% 
  right_join(counties, by=c("region", "subregion")) %>%
  ggplot(aes(x = long, y = lat, fill = choropleth), color = 'gray') +
  geom_polygon(aes(group=group)) +
  ggtitle("Choropleth Deaths by County")

choro
```


8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.

```{r, message=FALSE}
summer <- acc %>%  
  filter(MONTH %in% c(6, 7, 8)) %>% 
  group_by(STATE) 
sum <- summer %>% summarize(count=nrow(summer))
sum$season <- 'summer'

winter <- acc %>%  
  filter(MONTH %in% c(12, 1, 2)) %>% 
  group_by(STATE) 
win <-winter %>% summarize(count=nrow(winter))
win$season <- 'winter'

season <- merge(sum, win, all=TRUE)

ggplot(season ,aes(x = season, y = count, fill = season)) +
  geom_bar(stat='identity') +
  ggtitle("Accdients by Season for all States")
```

Summer seems more dangerous than winter.